The resurgence of the linear optics interferometer --- recent advances \& applications

### 2017-09-27 paper.tex
Added QKD review references to LOQC subsection (Section 10.1)

Another approach to linear optics is to pack more features known as 
degrees of freedom onto a single photon, and to increase the degree 
of control over them. It is possible to manipulate degrees of
freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, 
bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM, new proposals 
for linear optical quantum computations (LOQC) have surfaced 
\cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. 
\sihui {Some of them support new functionalities like secure delegated quantum 
computations \cite{bib:Barz12,bib:Fisher14}}. Others do not promise 
universality in their processing, but still performs tasks such as sampling 
\cite{bib:Aaronson11}, \sihui{encrypted quantum walks \cite{bib:Rohde_qw12}}, 
and secure quantum computations \cite{bib:Tan16} that might be hard or 
impossible with a classical system.  Such methods open up new ways to engineer 
quantum states and quantum operations, and expand the toolbox of baseline 
resources we have for implementing quantum computations, which could prove 
crucial in the race to achieve quantum supremacy. \sihui{Quantum key distribution (QKD) fundamentally does not require interferometry, but sound QKD protocols using Mach-Zehnder interferometer have been proposed \cite{bib:Sasaki14}, and in some cases realized \cite{bib:Takesue15}.}

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in time-bin based architecture for linear optical networks, and photodectors in Sections \ref{sec:timebins} and \ref{sec:detector} respectively. Last, but not least, we review applications for linear optical interferometry in LOQC, {\sc bosonsampling}, and quantum metrology.

@@

Better single-photon sources (see Section \ref{sec:singlephotons}) are part of the solution. Improvement in control of their degrees of freedom have also led to more efficient designs of circuits \cite{bib:Zhou11,bib:Lanyon09} that enabled the optical implementation of quantum teleportation of the state of a single photon \cite{bib:Wang15}. In this way, three-qubit \cite{bib:Lanyon09,bib:Micuda13,bib:Patel16} and four-qubit \cite{bib:Starek16} gates have also been implemented using LOQC.

>>>>

Another approach to linear optics is to pack more features known as 
degrees of freedom onto a single photon, and to increase the degree 
of control over them. It is possible to manipulate degrees of
freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, 
bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM, new proposals 
for linear optical quantum computations (LOQC) have surfaced 
\cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. 
\sihui {Some of them support new functionalities like secure delegated quantum 
computations \cite{bib:Barz12,bib:Fisher14}}. Others do not promise 
universality in their processing, but still performs tasks such as sampling 
\cite{bib:Aaronson11}, \sihui{encrypted quantum walks \cite{bib:Rohde_qw12}}, 
and secure quantum computations \cite{bib:Tan16} that might be hard or 
impossible with a classical system.  Such methods open up new ways to engineer 
quantum states and quantum operations, and expand the toolbox of baseline 
resources we have for implementing quantum computations, which could prove 
crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in time-bin based architecture for linear optical networks, and photodectors in Sections \ref{sec:timebins} and \ref{sec:detector} respectively. Last, but not least, we review applications for linear optical interferometry in LOQC, {\sc bosonsampling}, \sihui{ quantum metrology, and quantum repeat networks}.

@@

Better single-photon sources (see Section \ref{sec:singlephotons}) are part of the solution. Improvement in control of their degrees of freedom have also led to more efficient designs of circuits \cite{bib:Zhou11,bib:Lanyon09} that enabled the optical implementation of quantum teleportation of the state of a single photon \cite{bib:Wang15}. In this way, three-qubit \cite{bib:Lanyon09,bib:Micuda13,bib:Patel16} and four-qubit \cite{bib:Starek16} gates have also been implemented using LOQC. \sihui{We also briefly mention here that single photons and waveplates are used in a related application, namely, quantum key distribution (QKD). While QKD fundamentally does not require interferometry \cite{bib:BennetBrassard84,bib:Ekert91}, some QKD protocols rely on the Mach-Zehnder \cite{Bennett92, bib:Hughes2000,bib:Sasaki14,bib:Takesue15} and Franson interferometers \cite{bib:Ali-Khan07,bib:Brougham13}. Practical quantum key distribution relying on photonic technology is reviewed in \cite{bib:Scarani09,bib:Hoi-Kwong14}.
}
